An Adolescent Mini-Neptune in the Kepler Field

### 2021-07-02 paper/authors.tex
wow, nature papers are short
add ttv vs local slope analysis
update ms

\correspondingauthor{L.\,G.\,Bouma}
\email{luke@astro.princeton.edu}

@@

\author[0000-0002-2792-134X]{J. L. Curtis}
\affiliation{Department of Astronomy, Columbia University, 550 West 120th Street, New York, NY 10027, USA}
\affiliation{Department of Astrophysics, American Museum of Natural History, Central Park West, New York, NY 10024, USA}

>>>>

\correspondingauthor{L.\,G.\,Bouma}
\email{bouma.luke@gmail.com}

@@

\author[0000-0002-2792-134X]{J. L. Curtis}
\affiliation{Department of Astronomy, Columbia University, 550 West 120th Street, New York, NY 10027, USA}
\affiliation{Department of Astrophysics, American Museum of Natural History, New York, NY 10024, USA}

\author[0000-0003-1298-9699]{K. Masuda}
\affiliation{Department of Earth and Space Science, Osaka University, Osaka 560-0043, Japan}

### 2021-07-02 paper/ms.tex
update ms

While thousands of exoplanets have been discovered orbiting nearby
stars, the vast majority of them are several billion years old.  This
makes it difficult to test theories for the origins of the
different families of planets, since many expected evolutionary
processes operate on timescales of less than 100 million years.

For instance, the ``mini-Neptune'' planets, thought to be made of
molten rocky cores and extended atmospheric envelopes of hydrogen and
helium, are expected to shrink in size by factors of several over
their first $10^8$ years \citep{owen_constraining_2020}.
Specifically, they start with sizes of 4--12\,$R_\oplus$ at the time
of disk dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
2--4\,$R_\oplus$ by 10$^8$ years.  This change is driven by a
combination of stellar irradiation and internal heat that powers an
outflow, which eventually depletes or entirely strips the envelope
from the rocky core \citep{Owen_Wu_2013,gupta_sculpting_2019}.
Discovering young planets and measuring their compositions and
atmospheric outflows are key steps toward testing this paradigm, which
is often invoked to explain the observed radius distribution of both
mature exoplanets \citep{Fulton_et_al_2017}, as well as younger
planets \citep{bouma_cluster_2020}.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{TEMP_Prot_vs_BpmRp.png}
			\includegraphics[width=0.49\textwidth]{TEMP_hrd.png}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf The \cn\ is $35\pm15$\,Myr old.} {\it Left:} TESS and
    Kepler rotations periods and dereddened Gaia colors, with the
    Pleiades \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
    \citep[650\,Myr;][]{douglas_poking_2017} shown for reference.
    Stars of a given color (mass) move up through time due to magnetic
    braking.  The \cn\ has a 40 Myr gyrochronal age, as calibrated
    against the LDB-age of IC\,2602.  {\it Right:} HR diagram of the
    same comparison. {\bf todo: caption improve}.
   \label{fig:age}
	}
\end{figure*}

% \begin{figure*}[t]
% 	\begin{center}
% 		\leavevmode
% 		\includegraphics[width=1\textwidth]{f4.pdf}
% 	\end{center}
% 	\vspace{-0.7cm}
% 	\caption{
% 		{\bf Zoomed view of the \sn\ light curve.}  
% 		The full Kepler dataset spans 1{,}437 days (3.9 years), sampled at
% 		30 minute cadence.  A one hundred day segment is shown here.  The {\it top panel} shows the
% 		\texttt{PDCSAP} median-subtracted flux in units of
% 		parts-per-thousand ($\times 10^{-3}$).  The dominant signal
% 		is induced by starspots.  The model for the
% 		stellar variability (orange line) is subtracted in the {\it
% 			bottom panel}, revealing the transits of \pn, as well as other
% 		deviations from the stellar variability model.  The 
% 		Figure Set available online shows the entire 3.9 years of
% 		available data.
% 		\label{fig:lightcurve}
% 	}
% \end{figure*}
% 
% \begin{figure*}[t]
% 	\begin{center}
% 		\leavevmode
% 		\includegraphics[width=1\textwidth]{f5.pdf}
% 	\end{center}
% 	\vspace{-0.7cm}
% 	\caption{
% 		{\bf Phase-folded transit of Kepler 1627b, with stellar
% 			variability removed.}  
%     The 1-$\sigma$ model uncertainties and the maximum {\it a
%     posteriori} model are shown as the faint purple band, and the dark
%     purple line.  Gray points are individual \texttt{PDCSAP} flux
%     measurements; black points are binned to 20 minutes, with a
%     representative error bar shown.
%     {\bf LGB FIXME: seems like the uncertainty is bigger than the RMS
%     in the out of transit...}
% 		\label{fig:phasefold}
% 	}
% \end{figure*}

@@

At the time of the main Kepler mission (2009--2013), only four open
clusters were known in the Kepler field: NGC\,6866, NGC\,6811,
NGC\,6819, and NGC\,6791, with ages spanning 0.7\,Gyr to 9\,Gyr
\citep{meibom_kepler_2011}.  Since that time, analyses of the
kinematic, photometric, and astrometric Gaia data have expanded our
knowledge of open cluster and moving group memberships \citep[{\it
e.g.},][]{cantatgaudin_gaia_2018,zari_3d_2018,kounkel_untangling_2019,Meingast2021}.
As part of our Cluster Difference Imaging Photometric Survey (CDIPS,
\citealt{bouma_cdipsI_2019}), we concatenated the available analyses
from the literature, which yielded a list of candidate young and
age-dated stars (see Section~\ref{app:targetlist}).

Comparing our young star list against the Kepler field yielded two
discoveries.  The first, to be discussed in an upcoming analysis
by J. Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr
open cluster Theia\,520 (UBC\,1).  At least six Kepler planets are
known in the cluster, including the Kepler-52 and Kepler-968 systems.
The second discovery, and the focus of this work, is that Kepler
observed a portion of the $\delta$\,Lyr cluster, shown in
Figure~\ref{fig:skychart}.

Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) was previosuly
reported to host a close-in mini-Neptune planet candidate,
Kepler\,1627Ab
\citep{2012ApJS..199...24T,thompson_planetary_2018}
Given the
planet's previous statistical validation by \citet{morton_false_2016},
we decided to examine the age of the star and its cluster more closely.
We collected the previously measured {\bf XXX} Kepler
rotation periods (CITEP MCQUILLAN2014), and measured {\bf YYY} new stellar
rotation periods for cluster members observed by TESS.
We also performed an isochronal analysis.
The results are shown in Figure~\ref{fig:age}.
From the ensemble perspective, which provides the strongest
constraints (SEE CITET SODERBLOM 2014),
the age of the star is $XX \pm YY$\,Myr.
Individual measurements of the star's rotation period and lithium
abundance corroborate this conclusion.

@@

If the \pn\ transit signal is created by a bonafide planet, this age
would make it the youngest planet yet found by the main Kepler
mission\footnote{The re-purposed K2 mission found two slightly younger
systems: K2-33b \citep{David_et_al_2017,Mann_K2_33b_2016} and V1298
Tau \citep{david_four_2019}}.  A portion of the light curve is shown
in Figure~\ref{fig:lightcurve}, and the phase-folded transit is shown
in Figure~\ref{fig:phasefold}.  The dominant signal in the
\texttt{PDCSAP} photometry is a quasi-period starspot signal with a
peak-to-peak amplitude that varies between roughly 2\% and 8\%,
depending on the degree of asymmetry in spot coverage between the
stellar hemispheres.  The starspot signal repeats every
$2.642\pm0.042$\,days, where we have adopted an uncertainty based on
the scatter observed between different windows (quarters) of the
time-series data.  The details of how we model this starspot signal to
derive the parameters of the planet are discussed in
Section~\ref{app:gptransit}.  The results, given in
Table~\ref{tab:posterior}, are consistent with a mini-Neptune sized
planet ($3.56\pm 0.04\,R_\oplus$) on a circular 7.2\,day orbit
($e<0.XX$), around
a G8V star ($0.91 \pm 0.05 R_\odot$). 
{\bf TODO quote the eccentricity limit}

Could the transit signal be produced by anything other than a planet
orbiting this near-solar analog?  \citet{morton_false_2016} considered
the false positive scenarios of a foreground
eclipsing binary, a hierarchical triple-star binary system, and a
background eclipsing binary.  \citet{morton_false_2016} reported the
latter scenario to be the only possible case, with a (model-dependent)
probability of $\approx10^{-5}$.  While this statistical validation is
reassuring, ``validated'' planets have been demonstrated to be false
positives \citep{shporer_three_2017}.  Most failures of the
statistical validation approach arise when attempting to separate the
smallest stars from identically-sized gas giant planets.  With a size
of $\approx 3.6\,R_\oplus$, Kepler\,1627b does not meet this
concern.
The fact that we detect a low-mass stellar companion in
archival Keck-NIRC2 Kp-band (2.12\,$\mu $m) imaging does not change
the conclusion: the companion contributes 1\% of the flux
observed by Kepler, which combined with the observed transit shape
cannot be invoked as an explanation for the observed data
(Section~\ref{app:companionstar}).

Two additional lines of evidence ultimately persuaded us in favor of
the planetary interpretation.  First, the $\approx$100 days of
1-minute short-cadence data enables a more precise measurement of the
transit impact parameter: {\bf todo numbers} it is constrained to be
$<0.62$ (3-$\sigma$) from the 1-minute cadenced data, compared to
$<0.XX$ (3-$\sigma$) from the 30-minute data.  The second, and
ultimately convincing line of evidence came from our transit timing
analysis.  {\bf todo: KM pls check numbers} We isolated each of the
144 observed transits to within $\pm4.5$\,hr of each transit, and
fitted each window with both {\it i)} a local second-order polynomial
and transit, and {\it ii)} a local linear trend.  We let the mid-time
of each transit float, and then calculated the residual between the
measured mid-time and that of a perfectly periodic orbit.  This
residual, the ``transit timing variation'' (TTV), is plotted in
Figure~\ref{fig:ttvslope} against the local linear slope.  Only a
handful of Kepler Objects of Interest have shown this correlation
\citep{holczer_time_2015}, which is most readily interpreted as a TTV
induced by unresolved starspot crossings \citep{mazeh_time_2015}.  
This is of course only possible if the planet is transiting the
primary star, which excludes the any of the aforementioned false
positive scenarios.
It also implies that the planet's orbit is prograde.  The latter
interpretation assumes that the dominant
photometric variability is induced by dark spots, rather than bright
faculae.  Given the observed transition of Sun-like stellar
variability from spot to faculae-dominated regimes, we expect this
inference to be reasonably secure
\citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.

@@

In short, Kepler\,1627Ab provides an interesting extremum in the ages
of the Kepler planets, and opens multiple avenues for study from the
ground.  Observations of its Doppler transit ($\Delta v_{\rm RM}
\approx 20$\ms) with HIRES or the upcoming KPF spectrograph could
yield a quantification of the stellar obliquity, and a means for
verifying the suggestion of a prograde order from transit timing
(CITEP,CITEP).  Similar transit spectroscopy aimed at detecting either
hydrogen or helium absorption could yield insight into the planet's
atmosphere, which is expected to be outflowing at a rate of {\bf
X.X}\,$M_\oplus\,{\rm yr}^{-1}$ (CITE).  A more challenging
measurement would be a measurement of the planet's mass, through
either detailed timing analyses of the Kepler data, or through a
high-cadence multi-color radial velocity campaign.  The mass would
yield constraints on both the planet's composition (``what is the
extent of the initial gaseous envelope for a typical mini-Neptune?''),
as well as its initial entropy (``hot'' or ``cold''? CITE).

The precision of the Kepler data itself may enable more detailed
analyses without even needing to acquire more data.  While our transit
timing study did not yield the detection of any clearly significant
planet-induced TTVs, the star's flaring may be suggestive of
star-planet interactions.  Appendix REFERENCE highlights an odd flare
timing coincidence that may be worthy of further examination.

Ultimately though, our age measurement for \pn\ was enabled
by identifying its connection to the \cn\ through the Gaia data.  The
Supplementary Table~\ref{tab:v05} enables similar crossmatches to
already known planets, as well as forthcoming planets from the TESS
mission (CITE RickerJATIS2014).  An alternative approach also likely
to bear fruit is to identify kinematic associations around exoplanet
host stars using the Gaia data, and to then verify these associations
with ancilliary rotation periods and spectroscopy (CITEP Tofflemire).

>>>>

While thousands of exoplanets have been discovered orbiting nearby
stars, the vast majority of them are several billion years old.  This
makes it difficult to test theories for the origins of the different
families of planets, since many evolutionary processes are expected to
operate on timescales of less than 100 million years.

For instance, the ``mini-Neptune'' planets, thought to be made of
molten rocky cores and extended atmospheric envelopes of hydrogen and
helium, are expected to shrink in size by factors of several over
their first $10^8$ years
\citep{owen_constraining_2020,kite_atmosphere_2020}.  Specifically,
they start with sizes of 4--12\,$R_\oplus$ at the time of disk
dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
2--4\,$R_\oplus$ by 10$^8$ years.  This change is driven by a
combination of stellar irradiation and internal heat that powers an
outflow, which can eventually deplete or entirely strip the envelope
from the rocky core \citep{Owen_Wu_2013,gupta_sculpting_2019}.
Discovering young planets and measuring their compositions and
atmospheric outflows are key steps toward testing this paradigm, which
is often invoked to explain the observed radius distribution of both
mature exoplanets \citep{Fulton_et_al_2017}, as well as younger
planets \citep{bouma_cluster_2020}.

% TODO: decide whether a paragraph about young planet obliquities is
% needed.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{TEMP_Prot_vs_BpmRp.png}
			\includegraphics[width=0.49\textwidth]{TEMP_hrd.png}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf The \cn\ is $35\pm15$\,Myr old.} {\it Left:} TESS and Kepler
    rotations periods and dereddened Gaia colors, with the Pleiades
    \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
    \citep[650\,Myr;][]{douglas_poking_2017} shown for reference.
    Stars of a given color (mass) move up through time due to magnetic
    braking.  The \cn\ has a 40 Myr gyrochronal age, as calibrated
    against the LDB-age of IC\,2602.  {\it Right:} HR diagram of the
    same comparison. {\bf todo: caption improve. fix figures.}.
   \label{fig:age}
	}
\end{figure*}

@@

At the time of the main Kepler mission (2009--2013), only four open
clusters were known in the Kepler field: NGC\,6866, NGC\,6811,
NGC\,6819, and NGC\,6791, with ages spanning 0.7\,Gyr to 9\,Gyr
\citep{meibom_kepler_2011}.  Since that time, analyses of the
kinematic, photometric, and astrometric Gaia data have expanded our
knowledge of open cluster and moving group memberships \citep[{\it
e.g.},][]{cantatgaudin_gaia_2018,zari_3d_2018,kounkel_untangling_2019,Meingast2021}.
As part of our Cluster Difference Imaging Photometric Survey (CDIPS,
\citealt{bouma_cdipsI_2019}), we concatenated the available analyses
from the literature, which yielded a list of candidate young and
age-dated stars (see Appendix~\ref{app:targetlist}).

Comparing our young star list against the Kepler field yielded two
discoveries.  The first, to be discussed in an upcoming analysis by J.
Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr open
cluster Theia\,520 (UBC\,1).  At least six Kepler planets are known in
the cluster, from the Kepler-52 and Kepler-968 systems
\citep{rowe_validation_2014,jontof-hutter_following_2021}.  The second
discovery, and the focus of this work, is that Kepler observed a
portion of the $\delta$\,Lyr cluster, shown in
Figure~\ref{fig:skychart}.

{\bf Jason Curtis todo}
The \cn\ contains
Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470), which 
was previously reported to host a close-in mini-Neptune candidate planet
\citep{2012ApJS..199...24T,thompson_planetary_2018}
Given the planet's previous statistical validation by
\citet{morton_false_2016}, we decided to examine the age of the star
and its cluster more closely.  We collected the previously measured
{\bf XXX} Kepler rotation periods (CITEP MCQUILLAN2014), and measured
{\bf YYY} new stellar rotation periods for cluster members observed by
TESS.  We also performed an isochronal analysis.  The results are
shown in Figure~\ref{fig:age}.  From the ensemble perspective, which
provides the strongest constraints (SEE CITET SODERBLOM 2014), the age
of the star is $XX \pm YY$\,Myr.  Individual measurements of the
star's rotation period and lithium abundance corroborate this
conclusion.

@@

If the \pn\ transit signal is created by a bonafide planet, this age
would make it the youngest planet yet found by the main Kepler
mission\footnote{The re-purposed K2 mission found two slightly younger
systems: K2-33b \citep{David_et_al_2017,Mann_K2_33b_2016} and V1298
Tau \citep{david_four_2019}}.  A portion of the light curve is shown
with the phase-folded transit
in Figure~\ref{fig:lc}.  The dominant signal in the
\texttt{PDCSAP} photometry is a quasi-period starspot signal with a
peak-to-peak amplitude that varies between roughly 2\% and 8\%,
depending on the degree of asymmetry in spot coverage between the
stellar hemispheres.  The starspot signal repeats every
$2.642\pm0.042$\,days, where we have adopted an uncertainty based on
the scatter observed between different portions (quarters) of the
time-series data.  The details of how we model this starspot signal to
derive the parameters of the planet are discussed in
Appendix~\ref{app:gptransit}.  The results, given in
Table~\ref{tab:posterior}, are consistent with a mini-Neptune sized
planet ($3.56\pm 0.04\,R_\oplus$) on a circular 7.2\,day orbit
($e<0.XX$), around a G8V star ($0.91 \pm 0.05 R_\odot$).  {\bf TODO
quote the eccentricity limit}

Could the transit signal be produced by anything other than a planet
orbiting this near-solar analog?  \citet{morton_false_2016} considered
the false positive scenarios of a foreground eclipsing binary, a
hierarchical triple-star binary system, and a background eclipsing
binary.  Based on the transit shape, \citet{morton_false_2016} found
the latter scenario to be the only possible case, with a
(model-dependent) probability of $\approx10^{-5}$.  While this
statistical validation is reassuring, ``validated'' planets have been
demonstrated to be false positives \citep{shporer_three_2017}.  Most
failures of the statistical validation approach arise when attempting
to separate the smallest stars from identically-sized gas giant
planets.  With a size of $\approx 3.6\,R_\oplus$, Kepler\,1627b does
not meet this concern.  The fact that we detect a low-mass stellar
companion in archival Keck-NIRC2 Kp-band (2.12\,$\mu $m) imaging does
not change the conclusion: the companion contributes 1\% of the flux
observed by Kepler, which combined with the observed transit shape
cannot be invoked as an explanation for the observed data
(Appendix~\ref{app:companionstar}).

{\bf todo: LGB to check numbers on full-dataset transit ecc and b} 
Two additional new lines of evidence support the planetary
interpretation.  First, the $\approx$100 days of 1-minute
short-cadence data reduces the cadence-induced smearing enables a more
precise measurement of the transit impact parameter
\citep{kipping_binning_2010}.  The planet transits with $b<0.YY$
(3-$\sigma$) from the 1-minute cadenced data, compared to $<0.XX$
(3-$\sigma$) from the 30-minute data.

{\bf todo: Kento to check numbers on per-transit analysis} 
The second line of evidence came from our transit timing analysis. We
isolated each of the 144 observed transits to within $\pm4.5$\,hr of
each transit, and fitted each window with both {\it i)} a local
second-order polynomial and transit, and {\it ii)} a local linear
trend.  We let the mid-time of each transit float, and then calculated
the residual between the measured mid-time and that of a perfectly
periodic orbit.  This residual, the ``transit timing variation''
(TTV), is plotted in Figure~\ref{fig:ttvslope} against the local
linear slope.  Only a handful of Kepler Objects of Interest have shown
this correlation \citep{holczer_time_2015}, which is most readily
interpreted as a TTV induced by unresolved starspot crossings
\citep{mazeh_time_2015}.  This is of course only possible if the
planet is transiting the primary star, which excludes the any of the
aforementioned false positive scenarios.  It also implies that the
planet's orbit is prograde.  The latter interpretation assumes that
the dominant photometric variability is induced by dark spots, rather
than bright faculae.  Given the observed transition of Sun-like
stellar variability from spot to faculae-dominated regimes, we expect
this inference to be reasonably secure
\citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.

@@

{\bf LGB TODO: ask James Owen what he thinks the mass loss rate is}
Kepler\,1627Ab therefore provides a new extremum in the ages of the
Kepler planets, and opens multiple avenues for study from the ground.
Observations of its Doppler transit are expected to induce a
Rossiter-McLaughlin anomaly with an amplitude of $\approx 20$\ms, and
could yield a quantification of the stellar obliquity, and a means for
verifying our transit-timing based suggestion of a prograde orbit.
Similar transit spectroscopy aimed at detecting either hydrogen or
helium absorption could yield insight into the planet's atmosphere,
which is expected to be outflowing at a rate of {\bf
X.X}\,$M_\oplus\,{\rm yr}^{-1}$ (CITE).  A more challenging
measurement would be a measurement of the planet's mass, through
either detailed timing analyses of the Kepler data, or through a
high-cadence multi-color radial velocity campaign.  The mass could
yield constraints on both the planet's composition as well as its
initial entropy \citep{owen_constraining_2020}.

The precision of the Kepler data itself may enable more detailed
analyses without even needing to acquire more data.  While our transit
timing study did not yield the detection of any clearly significant
planet-induced TTVs, the star's flaring may be suggestive of
star-planet interactions.  Appendix~\ref{app:flare} highlights an odd flare
timing coincidence that may be worthy of further examination by those
more knowledgeable than us.

Ultimately however, our age measurement for \pn\ was enabled by
identifying its connection to the \cn\ using the Gaia data.  The
Supplementary Table~\ref{tab:v05} enables similar crossmatches to
already known planets, as well as forthcoming planets from the TESS
mission \citep{ricker_transiting_2015,guerrero_tess_2021}.  An
alternative approach also likely to bear fruit is to identify
kinematic associations around exoplanet host stars using the Gaia
data, and to then verify these associations with ancilliary rotation
periods and spectroscopy \citep{tofflemire_tess_2021}.
